Road Accident Prediction in Time and Space

### 2019-04-22 paper/paper.tex
Paper: Add newline between each sentence

\begin{abstract}
Road accidents are an important issue of our modern societies, responsible for millions of deaths and injuries. In this paper we show how one can leverage open datasets from a city like Montreal, Canada, to create some accident prediction models, using state-of-the-art big data analytics methods in Python. Such models could then be used in the context of road accident prevention, but also to identify key factors that can lead to a road accident, and eventually, help to elaborate new policies. This study also explains how we dealt with the severe class imbalance issue of the accident prediction problem, a recurrent issue that touches many other scientific domains like medical diagnoses or fraud detection. In particular, we show how we implemented Balanced Random Forests, a variant of the Random Forests machine learning algorithm, into the Scala and Python APIs of Apache Spark big data framework.

@@

  In addition, many publications aim at predicting the severity of an accident using various information from the accident in order to understand what causes an accident to be fatal. Chong et al. \cite{Chong2005} used decision trees, neural network and an hybrid model using a decision tree and a neural network. They obtained thes best performances with the hybrid model which reached an accuracy of 90\% for the prediction of fatal injuries. They identified that the seatbelt usage, the light conditions and the alcohol usage of the driver are the most important features. Abellán et al. \cite{Abellan2013} also studied traffic accident severity by looking at the decision rules of a decision tree using a dataset of 1801 highway accidents. They found that the the type and cause of the accident the light condition, the sex of the driver and the weather were the most important features.

@@

\subsection{Datasets}
We used three public datasets provided by the city of Montreal and the government of Canada: 
\begin{itemize}
\item Montreal Vehicle Collision: This dataset provided by the city of Montreal contains all the road collisions reported by the police in Montreal from 2012 to 2018. For each accident, the dataset contains the date and localization of the accident, information on the number of injuries and death, the number of vehicles involved, and information on the road condition. We used only the localization and the date of the accident since we do not have the other piece of information when no accident happened. Another dataset with all vehicle collision in Canada is available but without the localization of the accident, therefore we restrained our analysis to the city of Montreal.
\item National Road Network: This dataset provided by the government of Canada contains the geometry of all roads in Canada. For each road segment, a few meta-data are given. For roads in Quebec, the speed limit and the surface type are not provided. The data was available in various format, we chose to use the Keyhole Markup Language, which is a standard of the Open Geospatial Consortium since 2008. This format is based on the Extensible Markup Language (XML), which make it easier to read using existing implementation of XML parsers. 
\item Data Historical Climate Data: This dataset provided by the government of Canada contains hourly weather information measured at different weather stations across Canada. For each station and every hour, the dataset provides the temperature, the humidity, the wind direction and speed, the visibility, the atmospheric pressure and observations of atmospheric phenomenon such as snow, fog, rain, etc.
\end{itemize}

@@

\subsection{Algorithms selected}
For this study, we chose to focus on tree-based algorithms because they have proven their efficiency compared to classical statistical methods and allow for easier interpretability than deep learning algorithms. In particular, we implemented the Balanced Random Forests (BRF) algorithm in Apache Spark to deal with class imbalance which is a prominent issue in road accident prediction. To validate the performance of our version, we run an experiment with a simple synthetic dataset to compare its performances with the standard implementation of Random Forest (RF) in Spark using random under-sampling of the majority class. As expected, Balanced Random Forest performed better. Balanced Random Forests in one of the two methods proposed by Chen, Liaw, and Breiman\cite{Chen2004} to deal with class imbalance when using Random Forest. Balanced Random Forests are like Random Forests, but with a difference during the bootstrapping phase, for each tree of the forest, a random under-sampling of the majority class is performed in order to obtain a balanced sample. Intuitively, Balance Random Forest is an adaptation of random undersampling of the majority class making use of the fact that Random Forests are an ensemble method. Random undersampling usually performs better than more advanced methods like SMOTE or NearMiss \cite{Branco2016}. Chen, Liaw and Breiman\cite{Chen2004} also proposed in the same paper, the Weighted Random Forest (WRF) which consists in giving more weight to the minority class at the building time of each tree and during the measure of the gain in impurity and during the class prediction of each terminal node. While neither method is clearly better than the other in terms of prediction power, we chose BRF because it is found to be more computationally efficient (because of the under-sampling). Interestingly, Wallace et al. \cite{Wallace2011} presents a theoretical analysis of the data imbalance problem and suggest to use methods like Balance Random Forest.

We decided to compare the performance of the Balanced Random Forests algorithm with the performances of one of the very efficient implementations of Gradient Boosted Trees: XGBoost which offers parameters to deal with data imbalance as well.

>>>>

\begin{abstract}
Road accidents are an important issue of our modern societies, responsible for millions of deaths and injuries.
In this paper we show how one can leverage open datasets from a city like Montreal, Canada, to create some accident prediction models, using state-of-the-art big data analytics methods.
Such models could then be used in the context of road accident prevention, but also to identify key factors that can lead to a road accident, and eventually, help to elaborate new policies.
This study also explains how we dealt with the severe class imbalance issue of the accident prediction problem, a recurrent issue that touches many other scientific domains like medical diagnoses or fraud detection.
In particular, we show how we implemented Balanced Random Forests, a variant of the Random Forests machine learning algorithm, into the Scala and Python APIs of Apache Spark big data framework.

@@

  In addition, many publications aim at predicting the severity of an accident using various information from the accident in order to understand what causes an accident to be fatal.
Chong et al.\cite{Chong2005} used decision trees, neural network and an hybrid model using a decision tree and a neural network.
They obtained thes best performances with the hybrid model which reached an accuracy of 90\% for the prediction of fatal injuries.
They identified that the seatbelt usage, the light conditions and the alcohol usage of the driver are the most important features.
Abellán et al. \cite{Abellan2013} also studied traffic accident severity by looking at the decision rules of a decision tree using a dataset of 1801 highway accidents.
They found that the the type and cause of the accident the light condition, the sex of the driver and the weather were the most important features.

@@

\subsection{Datasets}
We used three public datasets provided by the city of Montreal and the government of Canada: 
\begin{itemize}
\item Montreal Vehicle Collision: This dataset provided by the city of Montreal contains all the road collisions reported by the police in Montreal from 2012 to 2018.
For each accident, the dataset contains the date and localization of the accident, information on the number of injuries and death, the number of vehicles involved, and information on the road condition.
We used only the localization and the date of the accident since we do not have the other piece of information when no accident happened.
Another dataset with all vehicle collision in Canada is available but without the localization of the accident, therefore we restrained our analysis to the city of Montreal.
\item National Road Network: This dataset provided by the government of Canada contains the geometry of all roads in Canada.
For each road segment, a few meta-data are given.
For roads in Quebec, the speed limit and the surface type are not provided.
The data was available in various format, we chose to use the Keyhole Markup Language, which is a standard of the Open Geospatial Consortium since 2008.
This format is based on the Extensible Markup Language (XML), which make it easier to read using existing implementation of XML parsers.

\item Data Historical Climate Data: This dataset provided by the government of Canada contains hourly weather information measured at different weather stations across Canada.
For each station and every hour, the dataset provides the temperature, the humidity, the wind direction and speed, the visibility, the atmospheric pressure and observations of atmospheric phenomenon such as snow, fog, rain, etc.
\end{itemize}

@@

\subsection{Algorithms selected}
For this study, we chose to focus on tree-based algorithms because they have proven their efficiency compared to classical statistical methods and allow for easier interpretability than deep learning algorithms.
In particular, we implemented the Balanced Random Forests (BRF) algorithm in Apache Spark to deal with class imbalance which is a prominent issue in road accident prediction.
To validate the performance of our version, we run an experiment with a simple synthetic dataset to compare its performances with the standard implementation of Random Forest (RF) in Spark using random under-sampling of the majority class.
As expected, Balanced Random Forest performed better.
Balanced Random Forests in one of the two methods proposed by Chen, Liaw, and Breiman\cite{Chen2004} to deal with class imbalance when using Random Forest.
Balanced Random Forests are like Random Forests, but with a difference during the bootstrapping phase, for each tree of the forest, a random under-sampling of the majority class is performed in order to obtain a balanced sample.
Intuitively, Balance Random Forest is an adaptation of random undersampling of the majority class making use of the fact that Random Forests are an ensemble method.
Random undersampling usually performs better than more advanced methods like SMOTE or NearMiss \cite{Branco2016}.
Chen, Liaw and Breiman\cite{Chen2004} also proposed in the same paper, the Weighted Random Forest (WRF) which consists in giving more weight to the minority class at the building time of each tree and during the measure of the gain in impurity and during the class prediction of each terminal node.
While neither method is clearly better than the other in terms of prediction power, we chose BRF because it is found to be more computationally efficient (because of the under-sampling).
Interestingly, Wallace et al. \cite{Wallace2011} presents a theoretical analysis of the data imbalance problem and suggest to use methods like Balance Random Forest.
We decided to compare the performance of the Balanced Random Forests algorithm with the performances of one of the very efficient implementations of Gradient Boosted Trees: XGBoost which offers parameters to deal with data imbalance as well.
